The Heavy Photon Search Test Detector

### 2014-04-23 hps-testrun-nim.tex
Norman Graf's comments on 2nd draft.

\begin{abstract}
The  Heavy Photon Search (HPS), an experiment to search for a hidden sector photon in fixed target 
electroproduction, is preparing for installation at the Thomas Jefferson National Accelerator Facility 
(JLab) in the Fall of 2014.  As the first stage of this project, the HPS Test Run apparatus was 
constructed and operated in 2012 to demonstrate the experiment's technical feasibility and to  confirm 
that the trigger rates and occupancies are as expected. This paper describes the HPS Test Run 
apparatus and readout electronics and its performance. In this setting, a heavy photon can be identified 
as a narrow \ee{}  mass peak above the trident background or as a narrow mass peak with a decay 
vertex displaced from the production target, so charged particle tracking and vertexing are needed for 
its detection. In the HPS Test Run, charged particles are measured
with a compact forward silicon microstrip tracker inside a dipole magnet. Electromagnetic 
showers are detected in a PbW0$_{4}$ 
crystal calorimeter situated behind the magnet used to trigger the experiment and identify 
electrons and positrons. Both detectors are placed close to the 
beam line and split top-bottom. This arrangement  provides sensitivity to low-mass heavy photons,  
allows clear  passage of the unscattered beam, 
and avoids the spray of degraded electrons coming from the target. 
The discrimination between prompt and displaced \ee{} pairs requires the first layer of silicon 
sensors be placed only 10~cm downstream of the target. The expected signal is small, and the 
trident background huge, so the experiment requires very large statistics. Accordingly, the
HPS Test Run utilizes  high-rate readout and data acquisition electronics and a fast trigger to exploit 
the essentially 100\% duty cycle of the CEBAF accelerator.
\end{abstract}

@@

%% main text
\section{Introduction}
\label{introduction}
The heavy photon (\Aprime{}), aka a ``hidden sector'' or ``dark'' photon, is a particle with mass 
$10-1000$~MeV which couples weakly to electric charge by mixing with the Standard Model 
photon~\cite{Holdom:1985ag}. Consequently, it can be radiated by electrons and subsequently decay 
into \ee{} pairs, albeit at rates far below those of QED trident processes. Heavy photons have been 
suggested by numerous beyond Standard Model theories~\cite{Essig:2013lka}, to explain the 
discrepancy 
between theory and experiment of the muon's $g-2$ ~\cite{Pospelov:2008zw}, and as a possible 
explanation of recent astrophysical 
anomalies~\cite{Adriani:2008zr,FermiLAT:2011ab,Aguilar:2013qda}.  Heavy photons  couple directly to  
hidden sector particles with "dark" or "hidden sector" charge; these particles
could constitute all or some of the dark matter~\cite{ArkaniHamed:2008qn}. 
Current phenomenology highlights the $20-1000$~MeV/c$^{2}$ mass range, and 
suggests that  the coupling to electric charge, $\epsilon$e, has $\epsilon$ in the range of 
$10^{-3} -10^{-5}$. This range of parameters makes \Aprime{} searches viable in 
medium energy fixed target electroproduction~\cite{Bjorken:2009mm}, but requires large data sets and 
good mass resolution to identify a small mass peak above the copious QED background. At small 
couplings, \Aprime{} become long-lived, so detection of a displaced decay vertex can reject the prompt 
QED background and boost experimental sensitivity.

The HPS experiment~\cite{HPS_proposal_2010} uses both invariant mass and secondary
vertex signatures to search for \Aprime{}. It uses  a $\approx 1$~m long silicon tracking and vertexing detector
inside a dipole magnet to measure charged particle trajectories and a fast electromagnetic calorimeter 
just downstream of the magnet to provide a trigger and identify electrons. The experiment utilizes very 
high rate front end electronics and runs at high trigger rates (up to 50kHz), exploiting the 100\% duty 
cycle of the Thomas Jefferson National Accelerator Facility (JLab) CEBAF accelerator to accumulate 
the needed statistics.

The HPS Test Run, using a simplified version of the HPS apparatus, was proposed and 
approved at JLab as the first stage of HPS. 
Its purposes included demonstrating that the apparatus and data acquisition systems are 
technically feasible and  the trigger rates and occupancies to be encountered in electron-beam 
running are as simulated. Given dedicated running time with electron beams, the HPS Test Run 
apparatus is capable of searching for heavy photons in unexplored regions of parameter space.    
The key design criteria for HPS and the HPS Test Run are the same:
\begin{itemize}
\item large and uniform acceptance in the forward region close to the beam in order to catch boosted 
\Aprime{} decay products,
\item beam passage through the apparatus in vacuum, to eliminate direct interactions with the 
detector and minimize beam gas interactions, 
\item a flexible, redundant and efficient trigger selecting electron and positron pairs at rates up to 
50~kHz,
\item excellent track reconstruction efficiency for electrons and positrons,
\item good angular and momentum resolution to reconstruct invariant mass precisely,
\item excellent vertex resolution to discriminate displaced \Aprime{} decays from prompt QED 
backgrounds,
\item high rate electronics with excellent timing resolution to minimize out of time backgrounds,
\item data handling rates of 100~MB/s to permanent storage,
\item detector components that can survive and efficiently operate in a high radiation environment 
with local doses exceeding 100~Mrad.
\end{itemize}

@@

\section{Detector Overview}
\label{detector}
{\small
\begin{figure*}[t]
\begin{center}
    \includegraphics[width=0.8\textwidth]{figures/hps_testrun_rendering}
\caption{Rendering of the HPS Test Run apparatus installed on the beam line.}
\label{fig:testrundetector}
\end{center}
\end{figure*}
}
The HPS Test Run apparatus was designed to run in Hall~B at JLab using the CEBAF electron beam, a 
499~MHz beam, at energies between 2.2 and 6.6~GeV and currents between 200 and 600~nA.  
The overall design of the experiment follows from the kinematics of \Aprime{} production which 
typically results in a final state particle within a few degrees of the incoming beam, especially at low 
$m_{\Aprime{}}$. Detectors must therefore be placed close to the beam. 
The intense electron beam enlarges downstream after multiple scattering in the target and electrons 
which have radiated in the target disperse horizontally in the field of the analyzing magnet. Together 
they constitute a "wall of flame" which must be completely avoided. Accordingly, 
the apparatus is split vertically to avoid a "dead zone", the region within $\pm 15$~mrad of the beam 
plane. In addition, the beam is transported in vacuum through the tracker to minimize beam-gas 
interaction backgrounds. Even with these precautions, the occupancies of sensors near the beam 
plane are high, dominated by the
multiple Coulomb scattering of the primary beam, so high rate detectors, a fast trigger, 
and excellent time tagging are required to minimize their impact.  
The trigger comes from a highly-segmented  lead-tungstate (PbWO$_{4}$) crystal calorimeter located just 
downstream of the dipole magnet. 
 %\begin{multicols}{2}
\begin{center}
\begin{table*}[t]
{\small
\caption{Overview of the coverage, segmentation and performance of the HPS Test Run detector. }
\begin{tabular}{lccccccc}
\hline 
System & Coverage & \# channels & ADC & Time resolution & \# layers & Segmentation & Performance \\
 & (mrad) &  & (bit) & (ns) &  &  &  \\
\hline
SVT & $15<\theta_{y} < 70$ & 12800 & 14 & $\approx 2$~ns & 5  & $\approx 120~\mu$m $r-\phi$ & $\sigma_{d0,y}  \approx 100~\mu$m \\
& (5 hits) &  &  &  & (stereo layers) & $\approx 6~\mu$m $z$ & $\sigma_{d0,x} \approx 300~\mu$m \\
& &  &  &  &  &  & $\sigma_{d0,z}\approx 1$~mm \\
\hline
ECal & $15<\theta_{y} < 60$ & 442 & 12 & 4~ns & 1 & $1.33\times1.33$~cm$^2$  & $\sigma(E)/E \approx 4.5\%$ \\ 
 &  &  &  &  &  & $1.6\times1.6$~cm$^2$  &  \\ 
\hline
\end{tabular}
\label{tab:detector-overview}
}
\end{table*}
\end{center}
%\end{multicols}
A rendering of the apparatus installed on the beam line is shown in 
Fig.~\ref{fig:testrundetector} and an overview of the coverage, segmentation and performance is 
given in Tab.~\ref{tab:detector-overview}.

The silicon vertex tracking and vertexing detector for HPS Test Run, or SVT, resides in a vacuum 
chamber inside the Pair Spectrometer (PS) dipole magnet in Hall B at JLab. The magnetic field 
strength was 0.5~T oriented vertically throughout the run. The SVT has five 
measurement stations, or ``layers,'' beginning 10 cm downstream of the target. Each layer 
comprises a pair of closely-spaced silicon microstrip sensors responsible for measuring a single 
coordinate, or ``view''. Introduction of a small (50 or 100~mrad) stereo angle between the two 
sensors of each layer provides three-dimensional tracking and vertexing throughout the acceptance 
of the detector. In order to accommodate the dead zone, the 
SVT is built in two halves that are approximately mirror reflections of one another about the plane of the 
nominal electron beam.  Each layer in one half is supported on a common support plate with 
independent cooling and readout.

@@

 Data was taken on three different converter thicknesses with photon fluxes between 0.4 and 
 $1.3\times10^8$/s at photon energies between 0.55 and 5.5 GeV produced by a 30 to 90 nA electron 
 beam. Data was measured for both polarities of the PS dipole magnet.'
\begin{table}[]
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickn. & Duration &  $e^-$ on radiator \\
 (\%$X_0$) & (s) & ($\mu$C)    \\   
\hline
%\hline
1.6   & 911 &   24.4 \\ %24385.9     \\ %27 nA
%\hline
0.18   & 2640 &   193.5 \\ % 193508.9  \\  %73 nA
%\hline
0.45  & 2149 &     140.7 \\ %  140709.9  \\ %65.5 nA
%\hline
0    & 1279  &   88.1 \\ %88079.6  \\
\hline
\end{tabular}
}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
\end{center}
\end{table}
The photon beam line during the HPS Test Run produced a relatively large number of \ee{} pairs 
originating upstream of the converter position. This contribution was measured during data taking 
with ``empty'' converter runs, i.e. removing the converter but with all other conditions the same. 
The runs taken during the time dedicated to HPS Test Run is summarized in Tab.~\ref{tab:currents}.

@@

The Silicon Vertex Tracker (SVT) enables efficient reconstruction of charged particles and precision 
determination of their trajectories. These measurements allow \Aprime{} decays to be distinguished 
from background via simultaneous estimation of the invariant mass of \ee{} decay products and the 
position of decay vertexes downstream of the target.

The design of the SVT is primarily driven by direct physics requirements and constraints from the 
environment at the interaction region. The \Aprime{} decay products 
have momenta in the range of 1~GeV, so multiple scattering dominates mass and vertexing 
uncertainties for any possible material budget, so the SVT must minimize the amount of 
material in the tracking volume. The signal yield for long-lived \Aprime{} is very small, so 
the rejection of prompt vertexes must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10~cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close to 1.5~mm from the center of the beam, 
defining the 15~mrad ``dead zone" mentioned previously, to maximize low-mass \Aprime{} acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone," the 
radiation dose approaches $10^{15}$~electrons/cm$^2$/month, or roughly 
$3 \times 10^{14}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, 
requiring the sensors to be actively cooled.
% and~\ref{fig:rates}.
%===================
%\begin{figure}[]
%\includegraphics[width=7cm]{figures/radiation}
%\label{fig:radiation}
%\caption{{\small Simulation of the electron fluence 10~cm downstream of the target in the active region 
%of the SVT ($|Y|>0.15~$mm) looking downstream; electrons bend towards $+x$.}}
%\end{figure}
Meanwhile, very low-energy delta rays from 
beam-gas interactions multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions.

@@

\subsection{Components}
The sensors for the SVT are $p+$-on-$n$, single sided, AC coupled, polysilicon-biased microstrip 
sensors fabricated on $<100>$ silicon and have 30 (60)~$\mu$m sense (readout) pitch over their 
$4\times10$~cm$^2$ surface. This sensor technology was selected to match the requirement of 
a $<1$\% $X_0$ per layer, single-hit resolution better than 50~$\mu$m and tolerant of a radiation 
dose of approximately $1.5\times10^{14}$~\fluenceunit{} for a six month run. The sensors 
were purchased from the Hamamatsu Photonics Corporation for the cancelled 
Run 2b upgrade of the D\O~experiment~\cite{Denisov:2001aa} which satisfied the requirement that 
the technology must be mature and available within the time and budget constraints.

@@

The total SVT power consumption budget of about 50~W is removed by a water/glycol mixture 
circulated through a flexible manifold attached to the copper tubes in the cooling blocks. During the Test 
run the sensors where operated at around $23^{\circ}$C. The power consumption is dominated by five 
APV25 ASICs on each hybrid board consuming approximately 2~W, radiant heat load is less than 
0.5~W per sensor and leakage current is only significant in a small spot after irradiation.

@@

\subsection{Alignment}
The SVT was aligned using a combination of optical, laser and touch probe surveys at SLAC and JLab. 
The optical survey of individual modules with precision of a few $\mu$m are combined with a 
touch-probe survey of the overall SVT support structure, with 25-100~$\mu$m precision, to locate the 
silicon sensor layers with respect to the support plates and the mechanical survey balls on the base 
plate. After full assembly and installation of the SVT at JLab, a mechanical survey of the SVT base plate 
position inside the pair spectrometer vacuum chamber is used to determine the global position of the 
SVT with respect to CEBAF beam line. The resulting survey-based alignment has the position of the 
silicon sensors correct to within a few hundred microns measured from tracks in the HPS Test Run 
data. A more sophisticated global track-based alignment technique to reach final alignment precision 
well below 50~$\mu$m is being developed.

@@

The ECal module shown in Fig.~\ref{fig:ecal-module} is based on a tapered 160~mm long 
PbWO$_{4}$ crystal with a $13.3\times13.3$~mm$^2$ ($16\times16$~mm$^2$) front 
(rear) face wrapped in VM2000 multilayer polymer mirror film. 
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}
The scintillation light, approximately $3$ photoelectrons/MeV, is readout by a 
5$\times$5~mm$^2$ Hamamatsu S8664-55 Avalanche Photodiode (APD) with 75\% quantum 
efficiency glued to the rear face surface using MeltMount 1.7 thermal plastic adhesive. 
The low gain of APDs ($\sim 200$) was compensated with custom made preamplifier boards, which 
provide a factor of $2333$ amplification of the APD signal.

\subsection{Layout}
Similar to the SVT, the ECal is built in two separate halves that are mirror reflections of one another 
about the plane of the nominal electron beam to avoid interfering with the 15~mrad "dead zone". 
As shown in Fig.~\ref{fig:ecal}, the 221 modules in each half, supported by aluminum support frames, 
are arranged in rectangular formation with 5 layers and 46 crystals/layer except for the layer closest to 
the beam where 9 modules were removed to allow a larger opening for the outgoing electron and 
photon beams. Each half was enclosed in a temperature controlled box ($<1^{\circ}$~F stability and 
$<4^{\circ}$~F uniformity) to stabilize the crystal light yield and the operation of the APDs and its 
preamplifiers. 
Four printed circuit boards mounted on the backplane penetrated the enclosure and was used to 
supply the $\pm 5$~V operating voltage for the preamplifiers, 400~V bias voltage to the APDs, and to 
read out signals from the APDs. Each half of the ECal was divided into 12 bias voltage groups with a 
gain uniformity of about 20\%.

@@

\subsection{Signal readout}
\label{sec:fadc}
After a 2:1 signal splitter, $1/3$ of an amplified APD signal was fed to a 
single channel of a JLab flash ADC (FADC) board~\cite{fadc}. $2/3$ of the signal was sent to a 
discriminator module and then to a TDC for a timing measurement. 
%\begin{figure}[]
%\begin{center}
%{\small
%\includegraphics[width=4cm]{figures/FADC250_Photo_001.jpg}
%\caption{The JLab FADC250 VXS module.}
%\label{fig:fadc}
%}
%\end{center}
%\end{figure}
The FADC boards are high speed VXS modules digitizing up to 16 APD signals 
at 250~MHz and storing samples in 8~$\mu$s deep pipelines with 12-bit resolution. 
When a trigger is received, the part 
of the pipeline from 5 samples before and 30 after the signal crossed a programmable 
threshold (for the HPS Test Run this was set to $\approx70$~MeV) are 
summed and stored in a 17-bit register for readout. In addition a 4~ns resolution timestamp of the 
threshold crossing is reported in the readout for each pulse.
This scheme significantly compresses the data output of the FADC. During offline data analysis, a 
calibrated pedestal value is subtracted to obtain the actual summed energy.
Two 20-slot VXS crates with 14 (13) FADC boards were employed in the HPS Test Run to read out the 
top (bottom) half of the ECal.
In the HPS Test Run 385 out of 442 modules (87\%) were used in offline reconstruction, 39 modules 
were disabled or not read out (no FADC channel available, no APD bias voltage or masked out due to 
excessive noise) and 18 were masked offline due to noise. 
%\begin{figure}[]
%\begin{center}
%{\small
%\includegraphics[width=8cm]{figures/fadc_datapath_diagram.png}
%\caption{The two different readout modes of the FADC used in the HPS Test Run.}
%\label{fig:hps_trigger_data}
%}
%\end{center}
%\end{figure}

>>>>

\begin{abstract}
The  Heavy Photon Search (HPS), an experiment to search for a hidden sector photon in fixed target 
electroproduction, is preparing for installation at the Thomas Jefferson National Accelerator Facility 
(JLab) in the Fall of 2014.  As the first stage of this project, the HPS Test Run apparatus was 
constructed and operated in 2012 to demonstrate the experiment's technical feasibility and to  confirm 
that the trigger rates and occupancies are as expected. This paper describes the HPS Test Run 
apparatus and readout electronics and its performance. In this setting, a heavy photon can be identified 
as a narrow peak in the \ee{} invariant mass spectrum, above the trident background or as a narrow  
invariant mass peak with a decay vertex displaced from the production target, so charged particle 
tracking and vertexing are needed for its detection. 
In the HPS Test Run, charged particles are measured with a compact forward silicon microstrip tracker 
inside a dipole magnet. Electromagnetic showers are detected in a PbW0$_{4}$ 
crystal calorimeter situated behind the magnet used to trigger the experiment and identify 
electrons and positrons. Both detectors are placed close to the 
beam line and split top-bottom. This arrangement  provides sensitivity to low-mass heavy photons,  
allows clear  passage of the unscattered beam, 
and avoids the spray of degraded electrons coming from the target. 
The discrimination between prompt and displaced \ee{} pairs requires the first layer of silicon 
sensors be placed only 10~cm downstream of the target. The expected signal is small, and the 
trident background huge, so the experiment requires very large statistics. Accordingly, the
HPS Test Run utilizes  high-rate readout and data acquisition electronics and a fast trigger to exploit 
the essentially 100\% duty cycle of the CEBAF accelerator at JLab.
\end{abstract}

@@

%% main text
\section{Introduction}
\label{introduction}
The heavy photon (\Aprime{}), aka a ``hidden sector'' or ``dark'' photon, is a particle with mass 
$10-1000$~MeV/c$^{2}$ which couples weakly to electric charge by mixing with the Standard Model 
photon~\cite{Holdom:1985ag}. Consequently, it can be radiated by electrons and subsequently decay 
into \ee{} pairs, albeit at rates far below those of QED trident processes. Heavy photons have been 
suggested by numerous beyond Standard Model theories~\cite{Essig:2013lka}, to explain the 
discrepancy 
between theory and experiment of the muon's $g-2$ ~\cite{Pospelov:2008zw}, and as a possible 
explanation of recent astrophysical 
anomalies~\cite{Adriani:2008zr,FermiLAT:2011ab,Aguilar:2013qda}.  Heavy photons  couple directly to  
hidden sector particles with ``dark'' or ``hidden sector'' charge; these particles
could constitute all or some of the dark matter~\cite{ArkaniHamed:2008qn}. 
Current phenomenology highlights the $20-1000$~MeV/c$^{2}$ mass range, and 
suggests that  the coupling to electric charge, $\epsilon$e, has $\epsilon$ in the range of 
$10^{-3} -10^{-5}$. This range of parameters makes \Aprime{} searches viable in 
medium energy fixed target electroproduction~\cite{Bjorken:2009mm}, but requires large data sets and 
good mass resolution to identify a small mass peak above the copious QED background. At small 
couplings, \Aprime{} become long-lived, so detection of a displaced decay vertex can reject the prompt 
QED background and boost experimental sensitivity.

The HPS experiment~\cite{HPS_proposal_2010} uses both invariant mass and secondary
vertex signatures to search for \Aprime{}. It uses  a $\approx 1$~m long silicon tracking and vertexing 
detector
inside a dipole magnet to measure charged particle trajectories and a fast electromagnetic calorimeter 
just downstream of the magnet to provide a trigger and identify electrons. The experiment utilizes very 
high-rate front-end electronics and runs at high trigger rates (up to 50kHz), exploiting the 100\% duty 
cycle of the JLab CEBAF accelerator to accumulate the needed statistics.

The HPS Test Run, using a simplified version of the HPS apparatus, was proposed and 
approved at JLab as the first stage of HPS. 
Its purposes included demonstrating that the apparatus and data acquisition systems are 
technically feasible and  the trigger rates and occupancies to be encountered in electron-beam 
running are as simulated. Given dedicated running time with electron beams, the HPS Test Run 
apparatus is capable of searching for heavy photons in unexplored regions of parameter space.    
The key design criteria for HPS and the HPS Test Run are the same:
\begin{itemize}
\item large and uniform acceptance in the forward region close to the beam in order to catch boosted 
\Aprime{} decay products,
\item beam passage through the apparatus in vacuum, to eliminate direct interactions with the 
detector and minimize beam gas interactions, 
\item detector components that can survive and efficiently operate in a high radiation environment 
with local doses exceeding 100~Mrad.
\item high-rate electronics with excellent timing resolution to minimize out of time backgrounds,
\item a flexible, redundant and efficient trigger selecting electron and positron pairs at rates up to 
50~kHz,
\item data handling rates of 100~MB/s to permanent storage,
\item excellent track reconstruction efficiency for electrons and positrons,
\item good angular and momentum resolution to reconstruct invariant mass precisely,
\item excellent vertex resolution to discriminate displaced \Aprime{} decays from prompt QED 
backgrounds,
\end{itemize}

@@

\section{Detector Overview}
\label{detector}
{\small
\begin{figure*}[t]
\begin{center}
    \includegraphics[width=0.8\textwidth]{figures/hps_testrun_rendering}
\caption{Rendering of the HPS Test Run apparatus installed on the beam line.}
\label{fig:testrundetector}
\end{center}
\end{figure*}
}
The HPS Test Run apparatus was designed to run in Hall~B at JLab using the CEBAF 499MHz 
electron beam at energies between 2.2 and 6.6~GeV and currents between 200 and 600~nA.  
The overall design of the experiment follows from the kinematics of \Aprime{} production which 
typically results in a final state particle within a few degrees of the incoming beam, especially at low 
$m_{\Aprime{}}$. Detectors must therefore be placed close to the beam. 
The intense electron beam enlarges downstream after multiple scattering in the target and electrons 
which have radiated in the target disperse horizontally in the field of the analyzing magnet. Together 
they constitute a ``wall of flame'' which must be completely avoided. Accordingly, 
the apparatus is split vertically to avoid a ``dead zone'', the region within $\pm 15$~mrad of the beam 
plane. In addition, the beam is transported in vacuum through the tracker to minimize beam-gas 
interaction backgrounds. Even with these precautions, the occupancies of sensors near the beam 
plane are high, dominated by the
multiple Coulomb scattering of the primary beam, so high-rate detectors, a fast trigger, 
and excellent time tagging are required to minimize their impact.  
The trigger comes from a highly-segmented  lead-tungstate (PbWO$_{4}$) crystal calorimeter located just 
downstream of the dipole magnet. 
 %\begin{multicols}{2}
\begin{center}
\begin{table*}[t]
{\small
\caption{Overview of the coverage, segmentation and performance of the HPS Test Run detector. }
\begin{tabular}{lccccccc}
\hline 
System & Coverage & \# channels & ADC & Time resolution & \# layers & Segmentation & Performance \\
 & (mrad) &  & (bit) & (ns) &  &  &  \\
\hline
SVT & $15<\theta_{y} < 70$ & 12780 & 14 & $\approx 2$~ns & 5  & $\approx 120~\mu$m $r-\phi$ & $\sigma_{d0,y}  \approx 100~\mu$m \\
& (5 hits) &  &  &  & (stereo layers) & $\approx 6~\mu$m $z$ & $\sigma_{d0,x} \approx 300~\mu$m \\
& &  &  &  &  &  & $\sigma_{d0,z}\approx 1$~mm \\
\hline
ECal & $15<\theta_{y} < 60$ & 442 & 12 & 4~ns & 1 & $1.33\times1.33$~cm$^2$  & $\sigma(E)/E \approx 4.5\%$ \\ 
 &  &  &  &  &  & $1.6\times1.6$~cm$^2$  &  \\ 
\hline
\end{tabular}
\label{tab:detector-overview}
}
\end{table*}
\end{center}
%\end{multicols}
A rendering of the apparatus installed on the beam line is shown in 
Fig.~\ref{fig:testrundetector} and an overview of the coverage, segmentation and performance is 
given in Tab.~\ref{tab:detector-overview}.

The silicon vertex tracking and vertexing detector for the HPS Test Run, or SVT, resides in a vacuum 
chamber inside the Pair Spectrometer (PS) dipole magnet in Hall B at JLab. The magnetic field 
strength was 0.5~T oriented vertically throughout the run. The SVT has five 
measurement stations, or ``layers,'' beginning 10 cm downstream of the target. Each layer 
comprises a pair of closely-spaced silicon microstrip sensors responsible for measuring a single 
coordinate, or ``view''. Introduction of a small (50 or 100~mrad) stereo angle between the two 
sensors of each layer provides three-dimensional tracking and vertexing throughout the acceptance 
of the detector. In order to accommodate the dead zone, the 
SVT is built in two halves that are approximately mirror reflections of one another about the plane of the 
nominal electron beam.  Each layer in one half is supported on a common support plate with 
independent cooling and readout.

@@

 Data was taken on three different converter thicknesses with photon fluxes between 0.4 and 
 $1.3\times10^8$/s at photon energies between 0.55 and 5.5 GeV produced by a 30 to 90 nA electron 
 beam. Data was measured for both polarities of the PS dipole magnet.'
\begin{table}[]
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickn. & Duration &  $e^-$ on radiator \\
 (\%$X_0$) & (s) & ($\mu$C)    \\   
\hline
%\hline
0    & 1279  &   88.1 \\ %88079.6  \\
%\hline
0.18   & 2640 &   193.5 \\ % 193508.9  \\  %73 nA
%\hline
0.45  & 2149 &     140.7 \\ %  140709.9  \\ %65.5 nA
%\hline
1.6   & 911 &   24.4 \\ %24385.9     \\ %27 nA
\hline
\end{tabular}
}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
\end{center}
\end{table}
The photon beam line during the HPS Test Run produced a relatively large number of \ee{} pairs 
originating upstream of the converter position. This contribution was measured during data taking 
with ``empty'' converter runs, i.e. removing the converter but with all other conditions the same. 
The runs taken during the time dedicated to HPS Test Run is summarized in Tab.~\ref{tab:currents}.

@@

The Silicon Vertex Tracker (SVT) enables efficient reconstruction of charged particles and precise 
determination of their trajectories. These measurements allow \Aprime{} decays to be distinguished 
from background via simultaneous estimation of the invariant mass of \ee{} decay products and the 
position of decay vertexes downstream of the target.

The design of the SVT is primarily driven by direct physics requirements and constraints from the 
environment at the interaction region. The \Aprime{} decay products 
have momenta in the range of 1~GeV, so multiple scattering dominates mass and vertexing 
uncertainties for any possible material budget. The SVT must therefore minimize the amount of 
material in the tracking volume. The signal yield for long-lived \Aprime{} is very small, so 
the rejection of prompt vertices must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10~cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close as 1.5~mm from the center of the beam, 
defining the 15~mrad ``dead zone'' mentioned previously, to maximize low-mass \Aprime{} acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone'', the 
radiation dose approaches $10^{15}$~electrons/cm$^2$/month, or roughly 
$3 \times 10^{14}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, 
requiring the sensors to be actively cooled.
% and~\ref{fig:rates}.
%===================
%\begin{figure}[]
%\includegraphics[width=7cm]{figures/radiation}
%\label{fig:radiation}
%\caption{{\small Simulation of the electron fluence 10~cm downstream of the target in the active region 
%of the SVT ($|Y|>0.15~$mm) looking downstream; electrons bend towards $+x$.}}
%\end{figure}
Meanwhile, very low-energy delta rays from 
beam-gas interactions multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions.

@@

\subsection{Components}
The sensors for the SVT are $p+$-on-$n$, single sided, AC coupled, polysilicon-biased microstrip 
sensors fabricated on $<100>$ silicon and have 30 (60)~$\mu$m sense (readout) pitch over their 
$4\times10$~cm$^2$ surface. This sensor technology was selected to match the requirement of 
$<1$\% $X_0$ per layer, single-hit resolution better than 50~$\mu$m and tolerance of a radiation 
dose of approximately $1.5\times10^{14}$~\fluenceunit{} for a six month run. The sensors 
were purchased from the Hamamatsu Photonics Corporation for the cancelled 
Run 2b upgrade of the D\O~experiment~\cite{Denisov:2001aa} which satisfied the requirement that 
the technology must be mature and available within the time and budget constraints.

@@

The total SVT power consumption budget of about 50~W is removed by a water/glycol mixture 
circulated through a flexible manifold attached to the copper tubes in the cooling blocks. During the 
HPS Test Run the sensors were operated at around $23^{\circ}$~C. The power consumption is 
dominated by five 
APV25 ASICs on each hybrid board consuming approximately 2~W, radiant heat load is less than 
0.5~W per sensor and leakage current is only significant in a small spot after irradiation.

@@

\subsection{Alignment}
The SVT was aligned using a combination of optical, laser and touch probe surveys at SLAC and JLab. 
The optical survey of individual modules with a precision of a few $\mu$m was combined with a 
touch-probe survey of the overall SVT support structure, with 25-100~$\mu$m precision, to locate the 
silicon sensor layers with respect to the support plates and the mechanical survey balls on the base 
plate. After full assembly and installation of the SVT at JLab, a mechanical survey of the SVT base plate 
position inside the pair spectrometer vacuum chamber is used to determine the global position of the 
SVT with respect to the CEBAF beam line. The resulting survey-based alignment has the position of the 
silicon sensors correct to within a few hundred microns measured from tracks in the HPS Test Run 
data. A more sophisticated global track-based alignment technique to reach final alignment precision 
well below 50~$\mu$m is being developed.

@@

The ECal module shown in Fig.~\ref{fig:ecal-module} is based on a tapered 160~mm long 
PbWO$_{4}$ crystal with a $13.3\times13.3$~mm$^2$ ($16\times16$~mm$^2$) front 
(rear) face wrapped in VM2000 multilayer polymer mirror film. 
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{A schematic view of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}
The scintillation light, approximately $3$ photoelectrons/MeV, is read out by a 
5$\times$5~mm$^2$ Hamamatsu S8664-55 Avalanche Photodiode (APD) with 75\% quantum 
efficiency glued to the rear face surface using MeltMount 1.7 thermal plastic adhesive. 
The low gain of APDs ($\sim 200$) was compensated with custom-made preamplifier boards, which 
provide a factor of $2333$ amplification of the APD signal.

\subsection{Layout}
Similar to the SVT, the ECal is built in two separate halves that are mirror reflections of one another 
about the plane of the nominal electron beam to avoid interfering with the 15~mrad ``dead zone''. 
As shown in Fig.~\ref{fig:ecal}, the 221 modules in each half, supported by aluminum support frames, 
are arranged in rectangular formation with 5 layers and 46 crystals/layer except for the layer closest to 
the beam where 9 modules were removed to allow a larger opening for the outgoing electron and 
photon beams. Each half was enclosed in a temperature controlled box ($<1^{\circ}$~F stability and 
$<4^{\circ}$~F uniformity) to stabilize the crystal light yield and the operation of the APDs and its 
preamplifiers. 
Four printed circuit boards mounted on the backplane penetrated the enclosure and were used to 
supply the $\pm 5$~V operating voltage for the preamplifiers, 400~V bias voltage to the APDs, and to 
read out signals from the APDs. Each half of the ECal was divided into 12 bias voltage groups with a 
gain uniformity of about 20\%.

@@

\subsection{Signal readout}
\label{sec:fadc}
After a 2:1 signal splitter, $1/3$ of an amplified APD signal was fed to a 
single channel of a JLab flash ADC (FADC) board~\cite{fadc}. $2/3$ of the signal was sent to a 
discriminator module and then to a TDC for a timing measurement. 
%\begin{figure}[]
%\begin{center}
%{\small
%\includegraphics[width=4cm]{figures/FADC250_Photo_001.jpg}
%\caption{The JLab FADC250 VXS module.}
%\label{fig:fadc}
%}
%\end{center}
%\end{figure}
The FADC boards are high speed VXS modules digitizing up to 16 APD signals 
at 250~MHz and storing samples in 8~$\mu$s deep pipelines with 12-bit resolution. 
When a trigger is received, the part 
of the pipeline from 5 samples before and 30 after the signal which crossed a programmable 
threshold (for the HPS Test Run this was set to $\approx70$~MeV) are 
summed and stored in a 17-bit register for readout. In addition a 4~ns resolution timestamp of the 
threshold crossing is reported in the readout for each pulse.
This scheme significantly compresses the data output of the FADC. During offline data analysis, a 
calibrated pedestal value is subtracted to obtain the actual summed energy.
Two 20-slot VXS crates with 14 (13) FADC boards were employed in the HPS Test Run to read out the 
top (bottom) half of the ECal.
In the HPS Test Run 385 out of 442 modules (87\%) were used in offline reconstruction, 39 modules 
were disabled or not read out (no FADC channel available, no APD bias voltage or masked out due to 
excessive noise) and 18 were masked offline due to noise. 
%\begin{figure}[]
%\begin{center}
%{\small
%\includegraphics[width=8cm]{figures/fadc_datapath_diagram.png}
%\caption{The two different readout modes of the FADC used in the HPS Test Run.}
%\label{fig:hps_trigger_data}
%}
%\end{center}
%\end{figure}
